Drawing Math

### 2022-05-11 paper/Paper.tex
complexity

%% Custom Commands
\newcommand{\rationalPeriod}{{T^j}'}
\newcommand{\rationalPeriodTotal}{{T}'}
\newcommand{\statePeriod}{T^j}
\newcommand{\seedAngleJ}[2]{{\theta^{#2}_{#1}}}
\newcommand{\seedAngle}[1]{{\theta^j_{#1}}}
\newcommand{\seedAngleNotime}{{\theta}^j}
\newcommand{\state}[2]{{s_{#1}}^{#2}}
\newcommand{\stateNotime}[1]{s^{#1}}
\newcommand{\totalPeriod}{T}
\newcommand{\totalState}{\pmb{s}_i}
\newcommand{\totalStateNotime}{\pmb{s}}
\newcommand{\comb}{Comb}
\newcommand{\combWithState}[1]{Comb\left(s^1_{#1}, s^2_{#1}, ..., s^k_{#1}\right)}
\newcommand{\inclIndic}[2]{{\mathrm{incl}_{#1}^{#2}}}
\newcommand{\combSingleTerm}[1]{\prod_{j=1}^k \cos(\theta_j) ^ {\inclIndic{#1}{j}}}
\newcommand{\digSumPeriodic}{\mathrm{digSumPeriodic}^j}
\newcommand{\digSumPeriodicNoJ}{\mathrm{digSumPeriodic}}
\newcommand{\totalOverRationalFrac}{\frac{\totalPeriod}{\rationalPeriodTotal}}

@@

Now say that the turtle's movement is determined by $k$ seed parameters drawn from
the same set. Then, for some state space $\mathcal{S}$,
define $s_i^j \in \mathcal{S}$ to be some arbitrary
state associated with timestamp $i$ for the $j$th seed parameter where $j \in [k]$.
Also, define $\pmb{s}_i = (s_i^1, s_i^2, ..., s_i^k)$.
Next we will define a set of functions $SU^j: \mathcal{S} \rightarrow \mathcal{S}$
(for $S$tate $U$pdater) such that $s_{i + 1}^j = SU^j(s_i^j, i)$. Note that for
$j, a \in [m]$ where $j \neq a$, $s_{i+1}^j$ is determined solely by $s_i^j$ and $i$ and not
$s_i^a$.

@@

\subsection{Specifying the task ahead of us}
For our case, we consider $Comb, SU^j$ to all be memoryless (i.e.\ there output
is uniquely determined by the current input). So, we can simplify the overall question.
If, $\pmb{s}_i = \pmb{s}_{i + \ell T}$ for some $T \in \N$ and all $i \in \N$, then 
$\Delta p_{i} = \Delta p_{i + \ell T}$. So then,
$\Delta P_{i, i + \ell T} = \Delta P_{i, i + \ell' T}$ for all $\ell, \ell' \in \N$.
Thus, $T$ is a period of the change in position. We can thus
break down our problem into two parts:
\begin{enumerate}
  \item Finding the period, $T$, of the state $\pmb{s}$.
  \item Checking whether $\Delta P_{i, i + T} = 0$.
\end{enumerate}

\subsection{Some more restrictions on our problem}
We further restrict the problem by only considering the case where
$$
  Comb((., ., ., ., \theta^1), (., ., ., ., \theta^2), ..., (., ., ., ., \theta^k)) =
    \left(\combSingleTerm{1}, ..., \combSingleTerm{D}\right)
$$
where $\mathrm{incl}_d^j \in \set{0, 1}$ for $d \in [D]$ indicates whether to include a given
$x \in R$ determined by seed $j$ for position update in the $d$th dimension.

We will also let $\mathcal{S} = N^5$ where for $(n, d, b, \theta) \in \mathcal{S}$,
$n$ is the numerator of a rational in fraction form, $d$ is the denominator, 
$b$ is the base (i.e\. base 10, base 12, etc.), $i$ is the current time step,
and $\frac{2\theta}{b \pi}$ is an ``angle" associated with the state.
Moreover, we will only consider
$$
  SU^j(n, b, d, \theta) = (n, b, d, \theta + \mathrm{digit}(n, b, d, i) \mod b).
$$

where $\mathrm{digit}(n, b, d, i)$ gives us the $i$th digit of the decimal expansion of $\frac{n}{d}$
in base $b$.

@@

\subsubsection{Some intuition}
While the restrictions may seem arbitrary, they aptly match our original problem statement.
Take the three dimensional case with two seeds. The two seeds can be thought
of as the fraction which generates the updates to two angles, $\theta, \alpha$.
$\theta$ and $\alpha$ can then uniquely determine a direction in 3D space in which the 
turtle is pointing. 
So, at each time step $i + 1$, $\theta_{i + 1}$ equals to $\theta_i$, but rotated by
some fraction of $2 \pi$ determined by the $i$th digit of the decimal expansion of one fraction and its base.
% TODO: clear up.

@@

\section{Results}
\subsection{Finding period $T$}
\subsubsection{Finding the period of $\frac{n}{d}$}
We will first aim to find period $T$ of the state $\pmb{s}$.
For some $(n, b, d, \theta) \in \pmb{s}_0$, by \cite{MathOverflowFracPeriod},
we have that the period of the decimal expansion of $\frac{n}{d}$
can be determined by finding the smallest $\rationalPeriod$ such that
\begin{equation}
\label{eq:nd-period}
b ^ {\rationalPeriod} \equiv 1 \mod d.
\end{equation}
See, appendix \ref{AppendixPeriod} for more detail.
More generally though, any nontrivial $\rationalPeriod$ satisfying equation \ref{eq:nd-period}
will be a period of $\frac{n}{d}$.

@@

\begin{remark}[Polytime]
  TODO: remark about discrete log and similarity... (we think this is hard, but in QPP)...
\end{remark}

@@

Moreover, define function $\digSumPeriodic \in \Z_b$
such that
% TODO: define T within the func from another func
% TODO: bettr defn
\begin{equation}
  \digSumPeriodic = \sum_{i = i_0}^{i_0 + \rationalPeriodTotal} \mathrm{digit}(n, d, b, i)
\end{equation}

\begin{remark}[Finding the digit sum of a rational]
  aaaa
\end{remark}

\subsubsection{Finding a period of $\seedAngleNotime$} % TODO: redifine things in terms of theta...
For $(n^j, b^j, d^j, \theta_i^j) = s_i^j$,
recall that $\seedAngle{i + 1} = \seedAngle{i} + \mathrm{digit}(n, b, d, i) \mod b$.
So, after period $\rationalPeriodTotal$,
\begin{align*}
  \theta_{i + \rationalPeriodTotal} &= \left(\theta_{i} + \sum_{\ell = i}^{\rationalPeriodTotal + i} \mathrm{digit}(n, b, d, \ell)\right) \mod b\\
  &= \left(\theta_{i} +  \digSumPeriodic\right) \mod b\\
  &= \theta_{i} \mod b +  \digSumPeriodic \mod b.
\end{align*}
So, after $p$ periods of length $\rationalPeriodTotal$ where $p  \cdot \digSumPeriodic \equiv 0 \mod b$,
$$
\theta_{i + p\rationalPeriodTotal} = \theta_i.
$$
For simplicity, lets define
$$
  T^j = p\rationalPeriodTotal
$$
where $T^j$ is a period of the state for seed $j$.
% TODO: remove $i$ from the state and somehow have it be implicit....

@@

\subsection{Does it close? $\Delta P_{i, i + T} = 0$?}
So now that we know the period of $\totalStateNotime$, we can ask if
$\Delta P_{i, i + T} = 0$. % TODO: reference

@@

\subsubsection{Closing along a single dimension}
Finally, to see if a shape will close,
we just need to find if equation \eqref{eq:singleDelta0} holds true for all dimensions.
First, let $A_d = \set{j \mid j \in [k] \; ; \inclIndic{d}{j} = 1}$
We can then see that
\begin{align*}
&\sum_{i = 1}^T \combSingleTerm{d}\\
&=\sum_{i = 1}^T \prod_{j = 1}^k \left(\half \left(e^{\seedAngle{i} I} +  e^{-\seedAngle{i} I}\right)\right)^\inclIndic{d}{j}\\
&=2^{-|A|} \sum_{i = 1}^T \prod_{j \in A_d} \left(e^{\seedAngle{i} I} +  e^{-\seedAngle{i} I}\right)\\
\end{align*}

\newcommand{\periodFrac}{\frac{\totalPeriod}{\rationalPeriod}}
\newcommand{\periodFracRational}{\frac{\rationalPeriodTotal}{\rationalPeriod}}

We can then see that for $(n, b, d, \theta_i) \in \pmb{s}_i$ and $x, y \in \N$ where $y < \rationalPeriodTotal$,
\begin{align}
  \seedAngle{x \rationalPeriodTotal + y} &= \sum_{i = 0}^{x \rationalPeriodTotal + y} \mathrm{digit}(n, b, d, i) \notag \\
  &= \sum_{p = 0}^{(x- 1)\rationalPeriodTotal} \sum_{q=0}^{\rationalPeriodTotal - 1} \mathrm{digit}(n, b, d, p\rationalPeriodTotal + q)
    + \sum_{p = x \rationalPeriodTotal}^{x \rationalPeriodTotal + y} \mathrm{digit}(n, b, d, p) \notag \\
    % ARGS::....
  \label{eq:angleBreakdown}
  &= x \cdot \digSumPeriodic + \sum_{p = x \rationalPeriodTotal}^{\rationalPeriodTotal + y}\mathrm{digit}(n, b, d, p) \notag \\
  &= x \cdot \digSumPeriodic + \sum_{p = 0}^{y}\mathrm{digit}(n, b, d, p)
\end{align}
because $\mathrm{digit}(n, b, d, x\rationalPeriodTotal + \ell) = \mathrm{digit}(n, b, d, \ell)$ for any $\ell \in \N$
by definition of periodicity.

Observe that 
\begin{align*}
 &\prod_{j \in A_d} \left(e^{\seedAngle{p \rationalPeriodTotal + q} I} +  e^{-\seedAngle{p \rationalPeriodTotal + q} I}\right) \\
 =& \exp(\seedAngleJ{p \rationalPeriodTotal + q}{1} + \seedAngleJ{p \rationalPeriodTotal + q}{2} + ... + \seedAngleJ{p \rationalPeriodTotal + q}{d}) 
  + \exp(\seedAngleJ{p \rationalPeriodTotal + q}{1} - \seedAngleJ{p \rationalPeriodTotal + q}{2} + ... + \seedAngleJ{p \rationalPeriodTotal + q}{d}) + ... \\
  &+ \exp(\seedAngleJ{p \rationalPeriodTotal + q}{1} - \seedAngleJ{p \rationalPeriodTotal + q}{2} - ... - \seedAngleJ{p \rationalPeriodTotal + q}{d})
\end{align*}
which then equals
\begin{equation}
 \sum_{\beta \in \set{0,1}^{|A_d|}} 
    \exp\left(
        \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
        \seedAngle{p \rationalPeriodTotal + q}I
      \right).
\end{equation}

Then, 
\begin{align*}
 &2^{-|A|} \sum_{i = 1}^T \prod_{j \in A_d} \left(e^{\seedAngle{i} I} +  e^{-\seedAngle{i} I}\right)\\
 =& 2^{-|A|} \sum_{p= 0}^{\frac{\totalPeriod}{\rationalPeriodTotal} - 1} \sum_{q = 0}^{\rationalPeriodTotal - 1}\prod_{j \in A_d} \left(e^{\seedAngle{p \rationalPeriodTotal + q} I} +  e^{-\seedAngle{p \rationalPeriodTotal + q} I}\right)\\
 =&
  2^{-|A|} \sum_{p= 0}^{\frac{\totalPeriod}{\rationalPeriodTotal} - 1} \sum_{q = 0}^{\rationalPeriodTotal - 1}
 \sum_{\beta \in \set{0,1}^{|A_d|}} 
    \exp\left(
        \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
        \seedAngle{p \rationalPeriodTotal + q}I
      \right)\\
 =&
  2^{-|A|} 
 \sum_{\beta \in \set{0,1}^{|A_d|}} 
     \sum_{p= 0}^{\frac{\totalPeriod}{\rationalPeriodTotal} - 1} \sum_{q = 0}^{\rationalPeriodTotal - 1}
    \exp\left(
        \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
        \seedAngle{p \rationalPeriodTotal + q}I
      \right).
\end{align*}

\newcommand{\eqWTSInnerProd}{
  \sum_{p= 0}^{\frac{\totalPeriod}{\rationalPeriodTotal} - 1} \sum_{q = 0}^{\rationalPeriodTotal - 1}
    \exp\left(
        \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
        \seedAngle{p \rationalPeriodTotal + q}I
      \right)}

Then, lets fix some $\beta \in \set{0, 1}^{|A_d|}$, we will now show that 
\begin{equation}
\label{eq:wts01}  
  \eqWTSInnerProd
\end{equation}
always equals $0$ or $1$ for any choice of $\beta$.

Observe that 
\begin{align}
  \exp(\seedAngle{p \rationalPeriodTotal + q}I) &= \exp\left(p \cdot \digSumPeriodic +
    \sum_{\ell = p \rationalPeriodTotal}^{p \rationalPeriodTotal + q}\mathrm{digit}(n, b, d, \ell)\right) \tag{by equation \eqref{eq:angleBreakdown}} \\
    \label{eq:expBreakdown}
    &= \exp\left(p \cdot \digSumPeriodic\right) \exp\left(\sum_{\ell = 0}^{q}\mathrm{digit}(n, b, d, \ell)\right).
\end{align}

So then, by equation \eqref{eq:expBreakdown}, we get that
\begin{align}
&\exp\left(
    \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
    \seedAngle{p \rationalPeriodTotal + q}I = 
\right)\notag \\ 
\label{eq:decomposeExp}
=&
\exp\left(
    \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
    \cdot p \cdot
    \digSumPeriodic
\right)
\exp\left(
    \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
      \sum_{\ell = 0}^{q}\mathrm{digit}(n, b, d, \ell).
\right).
\end{align}

We then use \eqref{eq:decomposeExp} to show that \eqref{eq:wts01} equals
\begin{equation}
\label{eq:inner-outer-split}
  \sum_{p= 0}^{\frac{\totalPeriod}{\rationalPeriodTotal} - 1} \left[
    \exp\left(
      pI \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
      \digSumPeriodic
    \right)
  \left(
  \sum_{q = 0}^{\rationalPeriodTotal - 1}
    \exp\left(
        \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
          \sum_{\ell = 0}^{q}\mathrm{digit}(n, b, d, \ell)
    \right)\right)\right].
\end{equation}

\subsubsection*{Case 1}
Observe that the inner sum,
$
  \sum_{q = 0}^{\rationalPeriodTotal - 1}
  \exp\left(
      \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
        \sum_{\ell = 0}^{q}\mathrm{digit}(n, b, d, \ell)
  \right)
$,
is invariant for all $p$ in equitation \eqref{eq:inner-outer-split}. So then,
let the inner sum equal $C \in \Z$.  Moreover, note that 
\begin{align*}
  &\exp\left(
      pI \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
      \digSumPeriodic
    \right)\\
  =& \prod_{j \in A_d} \exp\left(-1^{\beta_{(j)}} pI \cdot \digSumPeriodic \right)
\end{align*}
and that 
\begin{equation*}
  \exp\left(-1^{\beta_{(j)}} pI \cdot \digSumPeriodic \right) = \exp\left(0\right) = 1
\end{equation*}
when $p = \frac{\totalPeriod}{\rationalPeriodTotal}$. Then, we can see that
$$
\prod_{j \in A_d} \exp\left(-1^{\beta_{(j)}} pI \cdot \digSumPeriodic \right) = 1
$$
when $p = \frac{\totalPeriod}{\rationalPeriodTotal}$.

And so, because $\digSumPeriodic$ is a constant, we can conclude that
\begin{equation*}
 \exp\left(
      I \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
      \digSumPeriodic
    \right) 
\end{equation*}
is a $\totalOverRationalFrac^th$ root of unity iff $\sum_{j \in A_d} -1 ^ {\beta_{(j)}} \neq 0$.

So, for $\sum_{j \in A_d} -1 ^ {\beta_{(j)}} \neq 0$, we have that
\begin{align*}
  \eqWTSInnerProd &= C \sum_{p= 0}^{\frac{\totalPeriod}{\rationalPeriodTotal} - 1}
    \exp\left(
      pI \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
      \digSumPeriodic
    \right) \\
    &= C \sum_{p = 0}^{\totalOverRationalFrac - 1}\exp\left(W_{\totalOverRationalFrac}^p\right) \\
    &= 0.
\end{align*}
where $W_{\totalOverRationalFrac}^p$ is the $\totalOverRationalFrac^{th}$ root of unity.

If $\sum_{j \in A_d} -1 ^ {\beta_{(j)}} = 0$, then 
\begin{align*}
  \eqWTSInnerProd &= C \sum_{p= 0}^{\frac{\totalPeriod}{\rationalPeriodTotal} - 1} \exp(0) \\
  &= 1.
\end{align*}

So, we have now shown that \eqref{eq:wts01} is always $0$ or $1$. In particular, we also have our first 
closure result. If, $\forall \beta \in \set{0, 1}^{|A_d|}$, $\sum_{j \in A_d} -1 ^ {\beta_{(j)}} \neq 0$,
\begin{align*}
  \sum_{i = 1}^T \combSingleTerm{d} &=  2^{-|A|} 
  \sum_{\beta \in \set{0,1}^{|A_d|}} 
      \sum_{p= 0}^{\frac{\totalPeriod}{\rationalPeriodTotal} - 1} \sum_{q = 0}^{\rationalPeriodTotal - 1}
     \exp\left(
         \sum_{j \in A_d} -1 ^ {\beta_{(j)}}
         \seedAngle{p \rationalPeriodTotal + q}I
       \right) \\
        &= 0
\end{align*}
for all $d \in D$.
So then, $\Delta P_{0, T} = 0$ whenever 
no degree $1$ multinomials with coefficients of $-1$ or $1$ and $|A_d|$ variables has a root of $(\digSumPeriodicNoJ^1, \digSumPeriodicNoJ^2, ..., \digSumPeriodicNoJ^{|A_d|})$.
In other words, if variables, $x_1 = \digSumPeriodicNoJ^1, x_2 = \digSumPeriodicNoJ^2, ...$, there does not exist a multinomial
of form $\pm x_1 \pm x_2 \pm ... \pm x_{|A_d|} = 0$.

% TODO: second part can be thought of as a constant for now... ie C2=
% Then, we can think of the inner first sum as a constant of sorts, (C1)
% And we know that pI(T/T') = 1 and that pI0 = 1. So, if C1 \neq 0, we get some roots of unity
% and then the whole sum is 0...
% If it does equal 0, you have to check if C2 = 0... don't really know how to do this...
% If C1 = 0 and C2 not equal to 0, then

% TODO: then think of notation for the "distributed" polynomial, then you can expand...
% / assignment of +- strings... oy vey

% TODO: no clue what is going on with C2... (maybe throw in some pictures...)

% TODO: overall goal, show this is in coNP...

% A bunch of sums and multiplies
% \cite{watrous2018theory}

@@

\appendix
\section{Finding the period of $\frac{n}{d}$}
\label{AppendixPeriod}
aaaaa

>>>>

%% Custom Commands
\newcommand{\rationalPeriod}{{T^j}'}
\newcommand{\rationalPeriodTotal}{{T}'}
\newcommand{\statePeriod}{T^j}
\newcommand{\rationalAngleJ}[2]{{\theta^{#2}_{#1}}}
\newcommand{\rationalAngle}[1]{{\theta^j_{#1}}}
\newcommand{\rationalAngleNotime}{{\theta}^j}
\newcommand{\state}[2]{{s_{#1}}^{#2}}
\newcommand{\stateNotime}[1]{s^{#1}}
\newcommand{\totalPeriod}{T}
\newcommand{\totalState}{\pmb{s}_i}
\newcommand{\totalStateNotime}{\pmb{s}}
\newcommand{\comb}{Comb}
\newcommand{\combWithState}[1]{Comb\left(s^1_{#1}, s^2_{#1}, ..., s^k_{#1}\right)}
\newcommand{\inclIndic}[2]{{\mathrm{incl}_{#1}^{#2}}}
\newcommand{\numbToAngle}{\frac{2\pi}{b^j}}
\newcommand{\numbToAngleNoJ}[1]{\frac{2\pi}{b^{#1}}}
\newcommand{\combSingleTermCos}[1]{\prod_{j=1}^k \cos\left(\numbToAngle \cdot \theta^j\right) ^ {\inclIndic{#1}{j}}}
\newcommand{\digSumPeriodic}{\sigma^j}
\newcommand{\digSumPeriodicNoJ}{{\sigma}}
\newcommand{\totalOverRationalFrac}{\frac{\totalPeriod}{\rationalPeriodTotal}}
\newcommand{\changeTotalOneDim}{\Delta P_{0, T}^d}
\newcommand{\digit}{\mathrm{digit}}
\newcommand{\commonBase}{\pmb{b}}
\newcommand{\numbToCommon}{\frac{\commonBase}{b^j}}
\newcommand{\commonToAngle}{\frac{2\pi}{\commonBase}}
\newcommand{\sinOrCos}{\phi^j}
\newcommand{\combSingleTerm}[1]{\prod_{j=1}^k \sinOrCos\left(\numbToAngle \cdot \theta^j\right) ^ {\inclIndic{#1}{j}}}

@@

Now say that the turtle's movement is determined by $k$ seed parameter drawn from
the same set. Then, for some state space $\mathcal{S}$,
define $s_i^j \in \mathcal{S}$ to be some arbitrary
state associated with timestamp $i$ for the $j$th seed parameter where $j \in [k]$.
Also, define $\pmb{s}_i = (s_i^1, s_i^2, ..., s_i^k)$.
Next we will define a set of functions $SU^j: \mathcal{S} \rightarrow \mathcal{S}$
(for $S$tate $U$pdater) such that $s_{i + 1}^j = SU^j(s_i^j, i)$. Note that for
$j, a \in [m]$ where $j \neq a$, $s_{i+1}^j$ is determined solely by $s_i^j$ and $i$ and not
$s_i^a$.

@@

\subsection{Specifying the task ahead of us}
For our case, we consider $Comb, SU_i^j$ to all be memoryless (i.e.\ there output
is uniquely determined by the current input). So, we can simplify the overall question.
If, $\pmb{s}_i = \pmb{s}_{i + \ell T}$ for some $T \in \N$ and all $i \in \N$, then 
$\Delta p_{i} = \Delta p_{i + \ell T}$. So then,
$\Delta P_{i, i + \ell T} = \Delta P_{i, i + \ell' T}$ for all $\ell, \ell' \in \N$.
Thus, $T$ is a period of the change in position. We can thus
break down our problem into two parts:
\begin{enumerate}
  \item Finding the period, $T$, of the state $\pmb{s}$.
  \item Checking whether $\Delta P_{i, i + T} = 0$.
\end{enumerate}

\subsection{Some more restrictions on our problem}
We further restrict the problem by only considering $\mathcal{S} = \N^4$ where for $(n, d, b, \theta) \in \mathcal{S}$,
$n$ is the numerator of a rational in fraction form, $d$ is the denominator, 
$b$ is the base (i.e.\ base 10, base 12, etc.),
and $\frac{2\theta}{b \pi}$ is an ``angle" associated with the state.

Then, let $\sinOrCos: \R \rightarrow \R$ equal $\cos$ or $\sin$.

Now, we will only consider
$$
  SU_i^j(n, b, d, \theta) = (n, b, d, \theta + \mathrm{digit}(n, b, d, i) \mod b).
$$
where $\mathrm{digit}(n, b, d, i)$ gives us the $i$th digit of the decimal expansion of $\frac{n}{d}$
in base $b$. For the sake of convenience, we will use the word ``rational parameter" 
instead of ``seed parameter" from here on out.

Moreover, we consider the case where
$$
  Comb((., ., ., ., \theta^1), (., ., ., ., \theta^2), ..., (., ., ., ., \theta^k)) =
    \left(\combSingleTerm{1}, ..., \combSingleTerm{D}\right)
$$
where $\mathrm{incl}_d^j \in \set{0, 1}$ for $d \in [D]$ indicates whether to include a given
$x \in R$ determined by rational parameter $j$ for position update in the $d$th dimension.

@@

Also, lets set 
$$\commonBase = \lcm{(n, b, d, \theta) \in \pmb{s}_0} b.$$
In other words,
$\commonBase$ can be thought of as a ``common base" among all rational parameters.

\subsubsection{Some intuition}
While the restrictions may seem arbitrary, they aptly match our original problem statement.
The original problem statement derives a spherical change in position based off of a rational number's
digit at a particular timestep. The polar change in position also has a fixed radius. Translating
from a polar to cartesian update then only requires products of $\sin$s and $\cos$s. See \cite{NDimSphericalCoord}
for more details.

Take the three dimensional case for instance. The turtle's update cartesian space
is given by 
\begin{align*}
  x &= \cos(\alpha)\\
  y &= \sin(\alpha) \cos(\beta)\\
  z &= \sin(\alpha) \sin(\beta)
\end{align*}
where $\alpha = \numbToAngleNoJ{1} \cdot \theta^1$ and 
$\beta = \numbToAngleNoJ{2} \cdot \theta^2$. We can thus see that our definition
of $\comb$ captures the three dimensional case.

% TODO: clear up.

@@

\section{Does it close?}
In understanding whether a set of given rationals, bases, and updated functions 
draw a closed shape in $D$ dimensional space, we first need to find the period of 
the update delta, $\Delta p_i$. We then know that the total update over a period will be repeated 
indefinitely. Consequently, we then seek to find the total change in position over a period. 
If the total change is $0$, the shape will close as the Turtle will end up at its starting point
after every period length. If the total update is nonzero, the Turtle will not draw a closed
shape.

\subsection{Finding period $T$}
\subsubsection{Finding the period of $\frac{n}{d}$}
We will first aim to find period $T$ of the state $\pmb{s}$.
For some $(n, b, d, \theta) \in \pmb{s}_0$, by \cite{MathOverflowFracPeriod},
we have that the period of the decimal expansion of $\frac{n}{d}$
can be determined by finding the smallest $\rationalPeriod$ such that
\begin{equation}
\label{eq:nd-period}
b ^ {\rationalPeriod} \equiv 1 \mod d.
\end{equation}
See, appendix \ref{AppendixPeriod} for more detail.
More generally though, any nontrivial $\rationalPeriod$ satisfying equation \ref{eq:nd-period}
will be a period of $\frac{n}{d}$.

@@

\begin{remark}[Complexity]
  Interestingly, period finding of rational numbers is intimately tied to the discrete
  log problem and factoring. For more information, check out \cite{MathOverflowFracPeriod}.
  This gives some intuition that this closure problem may not be in BPP (Bounded Error Polynomial Time),
  but may be in BQP (Bounded Error Quantum Polynomial Time) by \cite{Shor_1997}.
\end{remark}

@@

Moreover, define $\digSumPeriodic \in \Z_b$
such that
% TODO: define T within the func from another func
% TODO: bettr defn
\begin{equation}
  \digSumPeriodic = \sum_{i = i_0}^{i_0 + \rationalPeriodTotal} \mathrm{digit}(n, d, b, i).
\end{equation}
In other words, $\rationalPeriodTotal$ is the digital sum over one period.

\begin{remark}[Complexity]
  For $d > 2$, prime, and coprime to $b$, we can find $\digSumPeriodic$
  in polytime by multiplying $(b - 1) \cdot \frac{d-1}{2} \mod b$ \cite{OnDecSeq}. The authors are unsure
  as to the complexity of finding $\digSumPeriodic$ otherwise.
\end{remark}

\subsubsection{Finding a period of $\rationalAngleNotime$} % TODO: redifine things in terms of theta...
For $(n^j, b^j, d^j, \theta_i^j) = s_i^j$,
recall that $\rationalAngle{i + 1} = \rationalAngle{i} + \mathrm{digit}(n, b, d, i) \mod b$.
So, after period $\rationalPeriodTotal$,
\begin{align*}
  \theta_{i + \rationalPeriodTotal} &= \left(\theta_{i} + \sum_{\ell = i}^{\rationalPeriodTotal + i} \mathrm{digit}(n, b, d, \ell)\right) \mod b\\
  &= \left(\theta_{i} +  \digSumPeriodic\right) \mod b.
\end{align*}
So, after $p$ periods of length $\rationalPeriodTotal$ where $p  \cdot \digSumPeriodic \equiv 0 \mod b$,
$$
\theta_{i + p\rationalPeriodTotal} \equiv \theta_i + 0 \equiv \theta_i.
$$
For simplicity, lets define
$$
  T^j = p\rationalPeriodTotal
$$
where $T^j$ is a period of the state for rational parameter $j$.
% TODO: remove $i$ from the state and somehow have it be implicit....

@@

\subsection{Finding the change in position over a period}
So now that we know the period of $\totalStateNotime$, we can ask if
$\Delta P_{i, i + T} = 0$. % TODO: reference

@@

\begin{remark}[Complexity]
  The algorithm we provide in equation \eqref{eq:singleDelta0} runs in time exponential 
  in the size of the input assuming the Word RAM model. 
  The period for the rational generated from rational parameter $j$, $1 \leq \rationalPeriod \leq d^j$. Then,
  the period over all rationals generated from parameters is at most 
  $$
    \lcm{j \in [k]} \rationalPeriod \leq \prod_{j \in [k]} \rationalPeriod \leq \left(\max_{j \in [k]} \; d^j \right)^k.
  $$
  Then, 
$0 \leq \totalPeriod \leq \rationalPeriodTotal \cdot \lcm{j \in [k]} b^j \leq \rationalPeriodTotal \left(\max_{j \in [k]} b^j\right)^k$.
  And because evaluating the product in \eqref{eq:singleDelta0} takes $O(k)$ time,
  we have that the time for \eqref{eq:singleDelta0} is at most
  $$
    O\left(\left[\max_{j \in [k]} \; (b^j d^j)\right]^k \right).
  $$
  Because \eqref{eq:singleDelta0} must be computed for each dimension, the algorithm
  runs in 
  $$
    O\left(D \left[\max_{j \in [k]} \; (b^j d^j)\right]^k \right)
  $$
  time.
\end{remark}

@@

\appendix
\section{Proving Property 1}
\label{AppendixPeriod}
\include{PaperAppendix.tex}
